---
title: "Lab: The 12 Days of Christmas"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages for Functions
library(dplyr)
library(glue)
library(stringr)
library(purrr)
library(english)
```


### A Test Dataset

Run the code below to load up a dataset called `xmas` that contains the crucial 
information about the gifts in the song. We will use this dataset to test out
our functions as we work on them.

```{r}
xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")
```

---

## Step One:  Plurals

Note that the gifts are listed in singular: for example, on day five the
narrator receives "five golden rings", but the entry in the dataset for the gift
on day 5 simply says "ring".  

Using the skeleton of the `pluralize_gift()` function, complete the code so that 
the function takes a gift and returns the appropriate plural. 

```{r}
# Function that takes a noun and makes it plural
# Arguments -- gift -- A string or vector of strings
# Return -- A string or vector of strings with the pluralized words

pluralize_gift <- function(gift){
  
  #if the string ends with a y, replace the y with an ies
    if (str_sub(gift, -1, -1) == 'y'){
      
        gift <- gift %>% str_replace("y$", "ies")
        
        #if the string is "goose", replace the "oo" to "ee". I didnt generalize to replace "oo"s to "ee"s because it doesn't usually apply as the plural form of door is doors and the plural form of moose is moose.
      } else if (gift == "goose") {
        
        gift <- str_replace(gift, "oo", "ee")
       
        #in any other case, simply add an s to the end of the string. 
      } else {
        
        gift <- gift %>% str_replace("$", "s")
        
      }
  
    return(gift)
}

```

*Hint 1:* Note that the gifts on days six and nine have unusual pluralizations.
You may assume that in other datasets, there will be no special cases besides
these types.

*Hint 2:*  The following small example may be useful to you:

```{r, eval = TRUE}
my_names <- c("Kimberly", "Trini", "Jason", "Billy", "Zach", "Tommy")

my_names %>% 
  str_replace("y$", "ies")
```


**Important:**  You should **absolutely not** "hard-code" anything into this 
function; this function should work in general, not just for the items in the
12 Days of Christmas.  For example, the word "rings" should not appear anywhere
in the function.  I should be able to give it any gift and any day, and get back
the plural of that gift.


### Test Your Function

Try your function out on the gifts in the dataset.  
(Consider: is your function *vectorized*?  It does not have to be, but it may
make things simpler!)

```{r}

#pluralize_gift(xmas$Gift.Item)

#run the pluralize_gift function on each string in the Gift.Item column in xmas dataset
purrr::map_chr(xmas$Gift.Item, pluralize_gift)

```


---


## Step Two: Creating sentences

Write a function called `make_phrase` that takes as input the necessary information, 
and returns a phrase.  For example, 



```{r}
make_phrase <- function(num, num_word, item, verb, adjective, location) {
  

  ## Step 2: Pluralize the gift if needed
  ## Items with numbers larger than 1 should be pluralized
  #
  ## All other single items should have "a" in front
  
  # checks if the first letter of the item is a vowel
  vowel_start <- str_sub(item, 1, 1) %>% str_detect("^(aeiou)")
  
  # if number of items is more than 1 pluralize the item
  if (num > 1) {
    
    item <- pluralize_gift(item)
    
    # Single items starting with a vowel should have "an" in front
  } else if (vowel_start) {
  
      num_word <- str_replace(num_word , "one", "an")
    
      # Single items not starting with a vowel should have "a" in front
  } else {
    
      num_word <- str_replace(num_word , "one", "a")
    
  }
  
  
  glue::glue("{num_word} {adjective} {item} {verb} {location}") %>%
    str_squish() %>% 
    
    # Replace NAs with blank strings
    str_replace_all(" NA", "")
  
}
```

```{r}

make_phrase(num = 10, 
            num_word = "ten", 
            item = "lord", 
            verb = "a-leaping", 
            adjective = "", 
            location = "")

```

should return

```
"ten lords a-leaping"
```
# Test Your Function

Try your function out on the `xmas` data, by making a new variable containing
the daily phrases.

**Hint:** The `Day.in.Words` variable isn't quite what you want! You want `12`
to say `"twelve"` **not** `"twelfth"`. I suggest using the **english** package
to create numbered days and then use those in your `pmap_chr()`!

```{r}

xmas <- xmas %>%
  mutate(
    Day.in.Words = as.character(english(Day)),
    Full.Phrase = pmap_chr(list(Day, Day.in.Words, Gift.Item, Verb, Adjective, Location), 
                           make_phrase)
  )

```


---


## Step Three:  Iteration.

Write a function called `sing_line` that takes as input:

* A dataset

* A number indicating which day to sing about

* The name of a column in the dataset that contains the phrases for each day.

```{r}
library(crayon)
# Produces the string for one day of the song.
# Arguments -- dataset -- A data frame containing information about gifts
# Arguments -- line -- The number of the line for the day you want to sing about
# Arguments -- phrase_col -- The variable name for the column in the dataset that contains the gift phrases
# Return -- A string singing the line of the song with all gifts for the given day.

sing_line <- function(dataset, line, phrase_col){
  
  # Step 1: Setup the intro line
  
  # converts number to order. ex: 1 to first, 2 to second
  num_word <- english::ordinal(line)
 
  # printing this line with the ordinal value of the line provided.
  intro <- glue::glue("On the {num_word} day of Christmas, my true love sent to me,")
  # Step 2: Sing the gift phrases

   phrases <- dataset %>%
     
     # prints all the items within the days of the line value provided. (line value of 3 would print the first 3 days)
      filter(Day <= line) %>% 
     
     # get the phrases in vector form
      pull({{phrase_col}}) %>% 
     
     # reverse the vector in order for the items to be descending in number 
      rev()

     #paste connects all the phrases expcept that last one and connects them with a comma and a new line in between each phrase
    a <- paste(phrases[1:(line-1)], collapse = ",\n")
    
    # if line is greater than 1, connect all the previous phrases with the last phrase with a new line and a "and"
    if (line > 1){
      
      a <- str_c(a, phrases[line], sep = "\nand ")
      
    }
    # glue the intro with the phrases separated by a new line and return the result
    return (glue("{intro}\n{a}."))


}


```

For example,


```{r}
sing_line(xmas, 
         line = 2, 
          phrase_col = Full.Phrase)

```

should return

```
On the second day of Christmas, my true love sent to me,
two turtle doves 
and a partridge in a pear tree.
```

*Hint:*  The `{{phrase_col}}` part, which I provided for you, lets you use 
column names as arguments to a function.  Don't delete that line of code!

---

## Step Four: Use Your Functions!

Run the following code to test out your functions! The ouput should be the 
lyrics for the entire 12 Days of Christmas song. 

```{r}
xmas2 <- xmas %>%
  mutate(
    day.num = as.character(english::english(Day)), 
    Full.Phrase = pmap_chr(list(num = Day, 
                                num_word = day.num, 
                                item = Gift.Item, 
                                verb = Verb, 
                                adjective = Adjective, 
                                location = Location), 
                           make_phrase)
  )

map_chr(1:12, ~sing_line(xmas2, .x, Full.Phrase)) %>%
  str_c(collapse = "\n") %>%
  cat()

```


## Step Five: Show Me Your Work! 

#### Get it on GitHub

Once you are done with all the functions, make sure you have **pushed** your
work to GitHub.

Make sure your repository contains your completed RMarkdown document and
knitted HTML. 

- Your HTML should **only** show the code for your functions and output from 
testing / verifying that your functions perform as they should. 

__If you used GitHub to store your work:__

Submit the link to your GitHub repository in the Lab 8 assignment portal. 

__If you downloaded the compressed folder and did not use GitHub:__

Submit **both** your RMarkdown file and knitted HTML file

